Contextual recruitment of selective attention can be updated via changes in task-relevance

### 2019-07-17 manuscript_pretty.Rmd
working - incomplete

source("data-analysis/analysis_E1.R")
source("data-analysis/analysis_E2.R")
source("data-analysis/analysis_E3.R")
source("data-analysis/table1.R")
source("data-analysis/analysis_E1_Exploratory.R")
```

Selective attention is commonly investigated using interference paradigms such as the Stroop [-@stroop_studies_1935] and flanker [@eriksen_effects_1974] tasks, where participants identify a target while ignoring a response-congruent or -incongruent distractor. Performance is typically better on congruent versus incongruent trials and the difference---the congruency effect---taken as an index of attentional priorities. Large congruency effects are thought to reflect ineffective filtering of the distracting stimuli whereas small congruency effects are thought to reflect effective filtering. By probing factors that systematically alter congruency effects, we can then make inferences about processes that control attentional filtering. For example, manipulating the frequency of conflict via the proportion of congruent versus incongruent trials has shown to influence the size of the congruency effect. Typically, a high proportion congruent experiment produces large congruency effects, whereas a low proportion congruent experiment produces small congruency effects [@logan_when_1979; @lowe_selective_1982; @west_effects_1998]. This result is usually  explained as strategic control, where participants increase attentional control under high-conflict demands and relax attentional control under low-conflict demands [@logan_attention_1980; @logan_strategies_1984; @logan_when_1979; @lowe_selective_1982]. Recent work however, has demonstrated that attentional control is not only adjusted by top-down regulation, but can also be triggered automatically by environmental cues [@brosowsky_memory-guided_2018; @bugg_support_2012; @egner_creatures_2014; @fischer_predicting_2015; @king_priming_2012; @mayr_outsourcing_2007].

@@

One explanation for CSPC effects is that the repeated application of attentional priorities in a particular context creates an associative link in episodic memory between the attentional control procedures and the contextual information [@abrahamse_grounding_2016; @brosowsky_memory-guided_2018; @crump_learning_2016; @egner_creatures_2014]. Once the associative link is established, processing the context is assumed to trigger the retrieval of previous experiences, automatically reinstating associated attentional priorities. As a result, automatic memory-based retrieval allows attentional priorities to be rapidly adjusted in a context-appropriate fashion. How associations between attentional priorities and context cues are formed and used still remains unclear however. The current study addresses two issues.

First, how is attentional control resolved when multiple contextual features could be used as cues? Second, it remains unclear how  Historically, theories of automaticity and attentional control have emphasized the role of practice and likened the process to skill-learning ().

@@

@@

The primary question of interest was whether the task-relevance of context features associated with different levels of conflict would influence the size of the congruency effect for frequency unbiased items (see Table 1). To that end, mean correct RTs and mean error rates from frequency unbiased trials were submitted to a mixed analysis of variance (ANOVA) with task-relevant context (0% and 100% PC) and unbiased-item congruency (congruent and incongruent) as within-subject factors, and context-type (object, social, and social/non-repeating) as the between-subjects factor.

The results of the RT analysis revealed the critical, two-way interaction between task-relevant context and unbiased-item congruency to be significant, `r RT_ANOVA$Task_Relevant_Context_Congruency`, demonstrating smaller congruency effects when the context dimension associated with high conflict was made task-relevant. Additionally, the three-way interaction between the task-relevant context, unbiased-item congruency, and context-type was non-significant, `r RT_ANOVA$Condition_Task_Relevant_Context_Congruency`, showing no evidence for differences between the CSPC effects across context-type manipulations.

Similarly, the corresponding error analysis also resulted in a significant two-way interaction between the task-relevant context and unbiased-item congruency, `r ACC_ANOVA$Task_Relevant_Context_Congruency`, and a non-significant three-way interaction between task-relevant context, unbiased-item congruency, and context-type, `r ACC_ANOVA$Condition_Task_Relevant_Context_Congruency`. Therefore, the congruency effects in error rates were also smaller when the context dimension associated with high conflict was made task-relevant, corroborating the results of the reaction time analysis.

@@

```{r figure3, echo=FALSE, fig.width=3.6, fig.height=4, fig.cap='(ref:figure3)'}
#print figure 3
#source: data-analysis/analysis_E2.R
figure3
```

@@

@@

```{r figure4, echo=FALSE, fig.cap='(ref:figure4)', fig.width=3.6, fig.height=3.5}
#print figure 3
#source: data-analysis/analysis_E2.R
figure4
```

@@

>>>>

source("data-analysis/analysis_E1.R")
source("data-analysis/analysis_E2.R")
source("data-analysis/analysis_E3.R")
source("data-analysis/table1.R")
source("data-analysis/analysis_E1_Exploratory.R")
source("data-analysis/plot-power-simulation.R")
```

Selective attention is commonly investigated using interference paradigms like the Stroop [-@stroop_studies_1935] and flanker [@eriksen_effects_1974] tasks, where participants identify a target while ignoring a response-congruent or -incongruent distractor. Performance is typically better on congruent versus incongruent trials and the difference---the congruency effect---taken as an index of attentional priorities. Large congruency effects are thought to reflect ineffective filtering of the distracting stimuli whereas small congruency effects are thought to reflect effective filtering. By probing factors that systematically alter congruency effects, we can then make inferences about processes that control attentional filtering. For example, manipulating the frequency of conflict via the proportion of congruent versus incongruent trials has shown to influence the size of the congruency effect. Typically, a high proportion congruent experiment produces large congruency effects, whereas a low proportion congruent experiment produces small congruency effects [@logan_when_1979; @lowe_selective_1982; @west_effects_1998]. This result is usually  explained as strategic control, where participants increase attentional control under high-conflict demands and relax attentional control under low-conflict demands [@logan_attention_1980; @logan_strategies_1984; @logan_when_1979; @lowe_selective_1982]. Recent work however, has demonstrated that attentional control is not only adjusted by top-down regulation, but can also be triggered automatically by environmental cues [@brosowsky_memory-guided_2018; @bugg_support_2012; @egner_creatures_2014; @fischer_predicting_2015; @king_priming_2012; @mayr_outsourcing_2007].

@@

One explanation for CSPC effects is that the repeated application of attentional priorities in a particular context creates an associative link in episodic memory between the attentional control procedures and the contextual information [@abrahamse_grounding_2016; @brosowsky_memory-guided_2018; @crump_learning_2016; @egner_creatures_2014]. Once the associative link is established, processing the context is assumed to trigger the retrieval of previous experiences, automatically reinstating associated attentional priorities. As a result, automatic memory-based retrieval allows attentional priorities to be rapidly adjusted in a context-appropriate fashion. How associations between attentional priorities and context cues are formed and used still remains unclear however. The current study addresses two related issues.

First, how is attentional control resolved when multiple contextual features could be used as cues? Second, it remains unclear how

Historically, theories of automaticity and attentional control have emphasized the role of practice and likened the learning process to the automaticity of skill-learning ().

@@

## Data Analysis 
All analyses in this and the following experiments were performed in R [@r_software_2019] using a variety of R packages (). In particular, we used the afex and car packages [@r_afex_2019; @r_car_2019] to perform all the Null Hypothesis Significance Tests, the BayesFactor package to perform Bayesian analyses [@r_bayesfactor_2018], and the conflictPower package to perform the simulation-based sensitivity and power analyses [@r_conflictPower_2019]. All data and scripts can be found

We supplemented each Null Hypothesis Significance Test with a Bayes Factor analysis (Rouder, Speckman.. 2009). Using conventional frequentist testing it is not possible to quantify the evidence for a null effect. Bayes Factors, however, quantify the degree to which the observed data are compatible with either the alternative or null hypothesis (Dienes, 2014; Rouder et al., 2014). All Bayesian analyses were performed using the R package BayesFactor and Bayes Factors (BFs) were calculated using its default settings (Morey and Rouder, 2014). Following the guidelines recommended by Rouder et al. (2009)

@@

The primary question of interest was whether the task-relevance of context features associated with different levels of conflict would influence the size of the congruency effect for frequency unbiased items (see Table 1). To that end, congruency effects (incongruent minus congruent performance) from frequency unbiased trials were submitted to a mixed analysis of variance (ANOVA) with task-relevant context (0% and 100% PC) and context-type (object, social, and social/non-repeating) as the between-subjects factor.

The results of the RT analysis revealed the critical, two-way interaction between task-relevant context and unbiased-item congruency to be significant, `r RT_ANOVA$Task_Relevant_Context`, demonstrating smaller congruency effects when the context dimension associated with high conflict was made task-relevant. Additionally, the three-way interaction between the task-relevant context, unbiased-item congruency, and context-type was non-significant, `r RT_ANOVA$Condition_Task_Relevant_Context`, showing no evidence for differences between the CSPC effects across context-type manipulations.

Similarly, the corresponding error analysis also resulted in a significant two-way interaction between the task-relevant context and unbiased-item congruency, `r ACC_ANOVA$Task_Relevant_Context`, and a non-significant three-way interaction between task-relevant context, unbiased-item congruency, and context-type, `r ACC_ANOVA$Condition_Task_Relevant_Context`. Therefore, the congruency effects in error rates were also smaller when the context dimension associated with high conflict was made task-relevant, corroborating the results of the reaction time analysis.

@@

@@

```{r figure3, echo=FALSE, fig.width=3.25, fig.height=4, fig.cap='(ref:figure3)'}
#print figure 3
#source: data-analysis/analysis_E2.R
figure3
```

@@

```{r figure4, echo=FALSE, fig.cap='(ref:figure4)', fig.width=3.25, fig.height=3.5}
#print figure 3
#source: data-analysis/analysis_E2.R
figure4
```

@@

# Estimating reproducability by simulation-based sensitivity- and power-analyses

```{r figure5, echo=FALSE, fig.cap='(ref:figure5)', fig.width=6.875, fig.height= 5}
#print figure 3
#source: data-analysis/plot-power-simulation.R
figure5
```

(ref:figure5) Results from Experiment 3 showing congruency effects (incongruent - congruent) for frequency unbiased items in reaction times (left) and error rates (right) as a function of the task-relevant context (100% proportion congruent versus 0% proportion congruent). Task-relevance was manipulated through explicit instructions.
